\section{Pipeline as Contribution}\label{sec:pipeline}

\subsection{Architecture}
% Figure F1: pipeline diagram.

The workflow pipeline is an orchestration layer over two external binaries, ORCA's orca_2mkl \cite{ORCA}
and Multiwfn \cite{multiwfn}. The OMol25 release initially focused on energies and forces with a limited set of properties in their 
aselmdb metadata. They subsequently released the 4M electronic structure dataset hosted at Argonne National Lab with more complete coverage
of electronic properties. The expanded 4M segment includes a slew of additional information including  wavefunctions (as gbws), density matrices, 
and the raw outputs from the finalized ORCA calculations (.out, .engrad, orca_property.txt files, etc.). This 4M release is coupled to our post-processing 
pipeline to yield an even richer set of descriptors for the community to train models and evaluate beyond the standard gammut of energies and forces for 
machine-learned interatomic potentials(MLIPs). 

Our pipeline reads in the orca.out and wavefunction file from each OMol job folder to commence. 
The gbw, in particular, is the foundation for all downstream analysis. As Multiwfn has more complete compatability with .wfn/.wfx files, 
the first step the pipeline performs is a Molden conversion
through \texttt{orca\_2mkl} followed by EDF library loading inside Multiwfn so that ECPs
are considered in downstream analyses. Without this EDF loading, there are potential inconsistencies with how ECPs are treated \textcolor{red}{TODO: cite multiwfn manual where it says that.} 
Folders that fail the EDF load are flagged automatically (see Sec.~\ref{sec:dataset}) and re-queued.

The user selects a calculation tier through a single integer, \texttt{full\_set} $\in \{0,1,2\}$,
which cascades through the four post-DFT families implemented in
\texttt{qtaim\_gen/source/data/multiwfn.py}. Tier 0 is the cheapest baseline with 
four partial charge schemes (Hirshfeld, ADCH, CM5, Becke), one bonding scheme (fuzzy bond
order), full QTAIM analysis, and Becke/Hirshfeld fuzzy density with optional
spin variants. Tier 1 adds three more charge schemes (VDD, MBIS, CHELPG), IBSI bond order,
and ELF + MBIS fuzzy density. Tier 2 adds Bader (basin-integrated) charges, Laplacian bond
orders, and Laplacian and gradient-norm fuzzy densities. Tier 2 enables the more
expensive electrostatic potential geometry analysis (binned into our other group of calculations). 
The tiers are subsets of each other, this signifies that a user can "stack" calculations from 0 to 2
to incrementally add richer descriptors as they go. The tiered design is intended to give users a
choice of how much compute they want to invest in the post-processing step, while alloing for future expansions.
Table~\ref{tab:tiers} summarizes the per-family contents of each tier; the QTAIM critical-point
search is run at every tier and the ORCA-derived globals are always parsed.

\begin{table}[h]
\centering
\small
\caption{Per-family contents of each calculation tier (\texttt{full\_set}). Tiers are
strict supersets: tier $k$ runs everything in tiers $0, \dots, k$. Schemes shown in a column
are those \emph{added} at that tier. Source: \texttt{qtaim\_gen/source/data/multiwfn.py}.}
\label{tab:tiers}
\begin{tabular}{lp{3.6cm}p{3.0cm}p{3.4cm}}
\toprule
\textbf{Family} & \textbf{Tier 0 (baseline)} & \textbf{Tier 1 (+adds)} & \textbf{Tier 2 (+adds)} \\
\midrule
Partial charges & Hirshfeld, ADCH, CM5, Becke & VDD, MBIS, CHELPG & Bader (basin-integrated) \\
Bond orders & Fuzzy bond order & IBSI bond order & Laplacian bond order \\
Fuzzy density & Becke density, Hirshfeld density, Becke spin, Hirshfeld spin & ELF, MBIS density, MBIS spin & Laplacian $\rho$, gradient-norm $|\nabla \rho|$ \\
Other (scalar fields) & Geometry (ellipticity), ALIE & --- & ESP geometry \\
\midrule
\multicolumn{4}{p{12cm}}{\textit{Always-on (run at every tier):} full QTAIM critical-point search (atoms, bond CPs, ring CPs, cage CPs; $\rho$, $\nabla^2\rho$, ellipticity, energy density); ORCA \texttt{.out} parsed globals (energies, orbital eigenvalues, dipole, gradient, Mulliken / Loewdin / Mayer charges, SCF metadata, quality flags).} \\
\bottomrule
\end{tabular}
\end{table}

Every job folder carries its own state of completed calculations and whether it is being worked on.
A \texttt{.processing.lock} file with a stale-mtime threshold protects against conflicting work 
across distributed workers, and a per-folder \texttt{out\_files.zip} archives every Multiwfn .out 
stream so a folder can be reparsed or analyzed without re-running conversions or Multiwfn. 
Validation is structural rather than numeric: each tier defines the JSON keys that must be 
present, and the pre-validation pass treats a folder as done only if all expected keys are populated. 
ECP-failed folders are detected by scanning the archived ADCH/CM5 logs for 
the \texttt{Loading EDF library finished} banner.

The per-folder outputs are seven JSON files: \texttt{charge.json},
\texttt{bond.json}, \texttt{qtaim.json}, \texttt{fuzzy\_full.json}, \texttt{other.json},
\texttt{orca.json}, and \texttt{timings.json}. The first five are Multiwfn-derived and
indexed by method name. The full per-method schema for each tier is enumerated in
Sec.~\ref{sec:dataset}; we only sketch the families here. \texttt{orca.json} is parsed
directly from the ORCA \texttt{.out} file (energies, orbital eigenvalues, dipole, gradient,
SCF metadata, Mulliken/Loewdin/Mayer charges, and quality-filter fields) by a single-pass
enum state machine implemented in \texttt{qtaim\_gen/source/core/parse\_orca.py}. We chose
this over a general parser like \texttt{cclib} \cite{cclib} because the state machine
streams the file once, has zero runtime dependencies beyond the standard library, and only
extracts the fields we need. \texttt{timings.json} stores per-step wallclock seconds and 
is carried through to LMDB as a provenance record so every downstream artifact can be 
audited back to its compute cost. We do note that timings.jsons does not store number of cores, 
architectures, or other details for exact comparisons but is still useful for relative analysis. 

Once the desired tier is complete, the JSON outputs are converted into a set of typed LMDB
files via \texttt{json-to-lmdb}: one LMDB per descriptor family (geometry, charge, bond,
qtaim, fuzzy, other, orca, timings) sharing a common key derived from the relative path of
the job folder. The compact LMDB representation is portable and self-describing, so it is
directly usable by external graph models such as Chemprop \textcolor{red}{TODO: cite}  or any custom PyG/DGL pipeline.
For users of our companion package \texttt{qtaim\_embed}, the \texttt{generator-to-embed}
entrypoint composes the typed LMDBs into PyG heterographs through the \texttt{GeneralConverter}
class, with configurable bonding schemes (structural cutoff, fuzzy, IBSI, or QTAIM),
charge filter, fuzzy filter, and ORCA-globals filter. A second \texttt{QTAIMConverter} class
is retained specifically for QTAIM-only graphs, where bond edges are defined by QTAIM bond
critical points rather than by structural cutoffs. The resulting graphs expose 
atom-level, bond-level, and global-level targets simultaneously.

\paragraph{Foundation for downstream ML work.} This pipeline is intented to provide not just 
a means to automate high-throughput DFT analysis, but also to be a stable foundation for 
scalable downstream ML work. LMDBs were used throughout, with matching keys across all descriptor
families, to provide a rapid source for data factories in ML models. Researchers can leverage this 
breathe of available information to train models for partial charges, bond orders, teacher-student 
models for downstream tasks, foundational models, etc. With the converters for \texttt{qtaim\_embed}, 
we also provide a working method for training graph neural networks (GNNs) \textcolor{red}{TODO: cite}
including a diverse set of message-passing and global-readout layers, more detail on that in Sec.~\ref{sec:ml_postprocessing}. 

\paragraph{Figure F1 prompt (for image generation).}
The following prompt can be passed to a vector-capable image generator (Nano Banana,
Claude, or similar) to render Figure F1:
\begin{quote}\small
\textit{Generate a clean horizontal SVG flow diagram of a quantum chemistry post-processing
pipeline. Left to right, four stages connected by labeled arrows.
Stage 1 (input): a single block labeled ``OMol25 \texttt{.gbw}'' with a small sub-arrow into
``\texttt{orca\_2mkl} $\rightarrow$ Molden + EDF''.
Stage 2 (compute, drawn as a stack of three escalating boxes labeled
``Tier 0: 4 charges, fuzzy bond, QTAIM, baseline fuzzy density'',
``Tier 1: +VDD, +MBIS, +CHELPG, +IBSI, +ELF/MBIS fuzzy'',
``Tier 2: +Bader, +Laplacian bond, +Laplacian/$\nabla\rho$ fuzzy, ESP geometry''),
with two side annotations: a small lock icon labeled
``\texttt{.processing.lock} (distributed safety)'' and a zip icon labeled
``\texttt{out\_files.zip} (reparse without recompute)''.
Stage 3 (per-folder JSON outputs): seven small file icons in a row labeled
\texttt{charge.json}, \texttt{bond.json}, \texttt{qtaim.json}, \texttt{fuzzy\_full.json},
\texttt{other.json}, \texttt{orca.json}, \texttt{timings.json}.
Stage 4 (LMDB + graphs): a database cylinder labeled ``Typed LMDBs (one per family,
shared keys)'' followed by two output branches, one labeled ``Portable: Chemprop / custom
PyG'' and one labeled ``\texttt{qtaim\_embed} $\rightarrow$ DGL heterographs (atom / bond /
global tasks)''. Use a muted blue and gray palette, sans-serif labels, no decorative icons
beyond the stated ones.}
\end{quote}

\paragraph{Features flagged for trimming before camera-ready.} While auditing the codebase
to write this section, two threads of dead or near-dead capability surfaced. First,
\texttt{qtaim\_gen/source/core/parse\_critic.py} (Critic2 critical-point parser) is no longer
imported by any active workflow; the only remaining reference is an optional
\texttt{critic2\_cmd} field in \texttt{create\_files.py} that gates a legacy reaction
workflow. We intend to drop the Critic2 dependency entirely so that Multiwfn is the sole
QTAIM backend. Second, the older QTAIM-only job runners under \texttt{scripts/old/}
(\texttt{generator\_run.py} and the QTAIM-specific Parsl runner variants) predate the
multi-method full-runner family and can be removed without loss of capability. We retain
\texttt{QTAIMConverter} itself because it underpins the bond-classification task.

\subsection{Sharded Execution and LMDB Merge}
% Figure F4: sharding schematic.

LMDB is single-writer and exhibits lock contention under naive parallel writes, so
\texttt{json-to-lmdb} does not attempt to share a single output environment across workers.
Instead, each parallel job is given a shard index $i \in \{0, \dots, N-1\}$ and a total shard
count $N$, and partitions the input folder list by index modulo $N$. For the flat
\textit{root/category/subset/job} hierarchy this is a glob plus a modulo filter; for jagged
hierarchies such as OMol4M, where job folders sit at variable depth, the user supplies an
explicit \texttt{--folder\_list} file of absolute paths and the same modulo partitioning
applies to the line index. In both cases the LMDB key is
\texttt{relpath(folder, root\_dir).replace(os.sep, "\_\_")}, which is identical across every
descriptor family, so the per-family shards always share keys and joining is trivial.

Each shard writes into its own subdirectory as \texttt{shard\_\{i\}.lmdb}. When the last
shard finishes its work it drops a sentinel file (\texttt{shard\_N.done}); the merge step
waits on the full sentinel set and then concatenates per-family shards into a single
consolidated LMDB per family. The same shard/merge pattern is implemented at the converter
stage in \texttt{converter.py}, so graph construction can also be parallelized and joined
into one final per-vertical graph LMDB.

\subsection{ML-Ready Post-Processing}\label{sec:ml_postprocessing}

The LMDB stage is the boundary between the pipeline and downstream machine-learning use.
Three features sit on top of that boundary and are intended to make benchmark construction
turnkey rather than reinventable.

The first is \emph{merging}: after a sharded run, the per-shard LMDBs across all descriptor
families are merged into one LMDB per family with a single pass that reads each shard, copies
records, and verifies key uniqueness. Merge is the operation that produces the artifact a
benchmark consumer actually reads.

The second is \emph{multi-vertical splitting}. The \texttt{multi-vertical-merge} entrypoint
takes a pipeline JSON config that lists multiple verticals (e.g.\ SPICE, QM9, RMechDB,
OMol25 subsets) and produces per-vertical train/val/test graph LMDBs whose splits are
globally consistent: a molecular formula that appears in two verticals is assigned to the
same split partition in both, so a held-out composition is held out everywhere. The
implementation lives in \texttt{qtaim\_gen/source/utils/multi\_vertical.py} and runs in
three phases (Plan, Build, Scale).

The third is \emph{train-only scaler fitting}. Feature scalers (atomic, bond, and global)
are fit on the train split only and applied to validation and test, eliminating the
quiet-but-common leakage path where a scaler fit on the full pool implicitly tells the
model the test-set mean. When \texttt{multi-vertical-merge} is used, the scaler is fit on
the union of all train splits across verticals and then applied per-vertical, which is the
right behavior when the downstream model is trained on the full mixture.

Together these three features turn a raw LMDB dump into something a benchmarking team can
load directly: composition-consistent splits, per-vertical files, leakage-safe scalers, and
deterministic keys.

\subsection{HPC Portability}
% ALCF, NERSC, single-node configs. One paragraph each.

We exercised the pipeline on three production HPC centers, each with a different scheduler
and a different parallelism paradigm. ALCF Polaris uses PBS Pro and a controller-and-workers
Parsl topology: one login or batch node hosts the Parsl controller, which farms individual
job folders out to PBS-allocated worker nodes through the \texttt{PBSProProvider} executor in
\texttt{qtaim\_gen/source/utils/parsl\_configs.py}. NERSC Perlmutter uses Slurm and the same
topology with the \texttt{SlurmProvider} executor; the only meaningful change is the queue
name and the per-node thread budget. Both centers scaled cleanly because the controller node
sees only the work queue and the worker nodes see only their assigned folders.

LLNL Tuolumne uses Flux as its native scheduler, and we did not get the controller-and-workers
topology to behave reliably under Flux. The fallback used on Tuolumne was the single-node
paradigm: a Slurm or Flux allocation grants a node, and a local Parsl
\texttt{ThreadPoolExecutor} chews through the local job queue with no remote dispatch. This
is the same path used for laptop and workstation runs (\texttt{full-runner}), parameterized
by \texttt{n\_workers} and \texttt{n\_threads\_per\_job}. NERSC and ALCF can also be run in
this single-node mode if a user prefers not to manage a controller, at the cost of losing
multi-node scaling.

The submission scripts and Parsl configs for all three centers are checked in under
\texttt{qtaim\_gen/source/utils/parsl\_configs.py}; PBS, Slurm, and the single-node Flux
fallback are all minimal edits of the same template. Reproducing a run on a new center is
expected to require swapping the provider class and the queue/account fields, nothing more.

\subsection{Throughput and Parallel Scaling}
% Figures F1/F3: throughput and scaling curves.

We report a single, reproducible local scaling experiment. A controlled HPC sweep is
impractical on the paper timeline because the ALCF, NERSC, and Tuolumne allocations are
committed to dataset production, and an HPC throughput number is not reproducible by a
reader anyway. The local sweep is run on a fixed structure set drawn from the test fixtures
under \texttt{tests/test\_files/}, on a single workstation, with ORCA 6.0.0 and Multiwfn
3.8 pinned. We sweep the number of concurrent job-folder workers $\in \{1, 2, 4, 8\}$ at
each of the three calculation tiers (\texttt{full\_set} $= 0, 1, 2$), holding the per-job
thread budget at its single-job-optimal value so that contention, not single-job tuning,
is what changes between points.

The reportable quantity is wallclock cost per folder as a function of concurrent workers,
plotted as one curve per tier. To our knowledge there is no published characterization of
Multiwfn under intra-node concurrency: existing Multiwfn benchmarks address per-job memory
and per-method correctness, not how the binary behaves when several copies share a node.
Even a clean single-node curve is therefore a useful contribution, and it has the
additional virtue that any reader with a workstation can rerun it.

\subsection{Reproducibility}
% ORCA + Multiwfn version pinning, deterministic conventions,
% end-to-end smoke test (F6).

The pipeline pins no specific version of ORCA or Multiwfn at runtime; some compatibility 
across versions of either is expected though we have not exhaustively tested versions of each.
For example, we also tested Orca5\textcolor{red}{TODO: cite} wfn conversion where our code handles 
its treatment of ECPs and as well as Orca 6.0 and 6.1. In the case of Multiwfn, any version that 
follows the same menu structure should work though we only tested 3.9 and 3.8.
The dataset reported here was produced with ORCA 6.0.0, the same release used for OMol25 and Multiwfn 3.8. 
In our experience, the orca_2mkl version should match that of the gbws from input DFT calculations.

Determinism in the pipeline is structural rather than numeric: the LMDB key derivation, 
the per-tier set of expected JSON keys, and the parsing logic for both
Multiwfn and ORCA outputs are fixed by code and covered by the test suite under
\texttt{tests/}. Numeric reproducibility of the underlying QM calculation is the
responsibility of ORCA itself and is not something we modify.

The test suite serves as the end-to-end smoke test: \texttt{tests/test\_lmdb.py} and
\texttt{tests/test\_general\_converter\_filters.py} together exercise json-to-lmdb,
sharded conversion, merge, the GeneralConverter graph construction, and scaler fit/apply on
a small fixture set under \texttt{tests/test\_files/lmdb\_tests/}. Running \texttt{pytest -q}
on a fresh checkout is therefore a ground-truth check that a user's local toolchain produces
the same LMDB and graph artifacts we ship.
