\section{Evaluation Protocol}
\label{sec:eval}

% Keystone for the D\&B track. All counts anchored to the 2026-04-28 manifest
% (3,986,738 ok rows, 100\% read success).

\subsection{Tasks}
\label{sec:eval-tasks}

\begin{itemize}
  \item \textbf{T1 -- Per-atom partial charge regression.} Multi-target:
    ADCH, Lowdin, Mulliken, Hirshfeld, CM5 jointly per atom. Headline:
    per-method, per-element MAE and Pearson $r$ on the test split.
    Long-tail per-element table in appendix.
  \item \textbf{T2 -- BCP property regression.} Targets: $\rho$,
    $\nabla^2\rho$, ellipticity at every QTAIM bond critical point.
    Headline: per-element-pair MAE on $\rho$, Pearson $r$ per pair,
    stratified by bond type (single, double, aromatic, metal-ligand)
    inferred from QTAIM topology.
  \item \textbf{T3 -- Bond classification.} Binary: does a QTAIM BCP
    exist between an atom pair within $2\times$ sum-of-covalent-radii.
    Compared against Mayer-ORCA / Loewdin-ORCA / fuzzy agreement. Macro-F1 plus
    per-element-pair confusion stats.
  \item T3 sharpening (open): mine the disagreement subset (QTAIM-bonded
    vs distance-bonded) once \texttt{bond.lmdb} and \texttt{qtaim.lmdb}
    are populated end-to-end; that subset is the discriminating slice.
\end{itemize}

\subsection{Main Split}
\label{sec:eval-main-split}

\begin{itemize}
  \item Composition-ordered, deterministic: each \texttt{formula\_hill}
    hashed via BLAKE2b~\cite{10.1007/978-3-642-38980-1_8} into $N=10$
    buckets.
  \item Buckets 0-7 train, bucket 8 val, bucket 9 test. Holdouts
    (\S\ref{sec:eval-holdouts}) are excluded before bucket assignment.
  \item Realized post-holdout sizes: train 3{,}189{,}390 / val
    399{,}690 / test 415{,}913.
  \item Bucket-size variance under 5\% of the mean; per-bucket
    unique-formula count $\approx$112k each, confirming the partition
    balances composition and not row count.
  \item Implementation: \texttt{split\_descriptor\_lmdbs.py} runs
    per-vertical, sharded over verticals, race-free under parallel
    \texttt{xargs}; per-vertical \texttt{split\_assignment.json} is the
    durable record.
  \item Verticals are not sampled uniformly because the split is global
    by formula. A per-vertical post-split distribution table will be
    published in the appendix as a transparency artifact.
\end{itemize}

\subsection{Held-Out Evaluation Sets}
\label{sec:eval-holdouts}

The held-out sets are constructed independently of the main split and
materialized before any model is trained. Membership is decided from the
manifest, and from \texttt{bond.lmdb} for the two metal-ligand sets,
then removed from the train, val, and test buckets so the same structure
is never simultaneously evaluated as in-distribution test and as a
stress test. Overlaps between held-out sets are allowed (a structure can
be both H7-large and H8-high-charge), but each set is materialized once
and reported separately so the reader can attribute error to the axis
being stressed. None of these sets is used for hyperparameter selection
or early stopping; numbers are reported once at the end of training.

The headline generalization signal per set is the gap
$\Delta = \mathrm{metric}_{\mathrm{held\text{-}out}} -
\mathrm{metric}_{\mathrm{main\,test}}$, reported alongside the absolute
number so a reviewer can read the floor (test) and the stress response
($\Delta$) on the same line. A small $\Delta$ on H7 means size
extrapolates; a large $\Delta$ on H1 means TM-ligand chemistry is being
memorized at the pair level.

Three of the five sets (H1, H6, H7) inherit the hardest evaluation axes
from the original OMol25 paper directly: rare transition-metal-ligand
bond pairs, rare lanthanide-ligand bond pairs, and large-system
extrapolation. Where OMol25 publishes a hold-out construction, we lift
it unchanged onto our manifest rather than redesigning it, both for
comparability and because the OMol25 axes were already chosen to stress
chemically meaningful failure modes.

\paragraph{Sets and sizes.}
The five held-out sets together cover roughly 60{,}719 structures
($\sim$1.5\% of the dataset; overlaps allowed):

\begin{itemize}
  \item \textbf{H1 -- Metal-ligand pairs (15{,}030 structures).} Eleven
    (TM, partner) bond pairs sampled stratified across log-spaced
    frequency bands with shape $(3,2,2,4,0)$ and \texttt{seed=87}.
    Selection is driven by \emph{true} bonds (Mayer / fuzzy bond order
    $\geq 0.3$) extracted from \texttt{bond.lmdb}, not element
    co-occurrence in the formula; the full procedure is in
    Appendix~\ref{app:heldout-h1}. Tests whether the model can predict
    descriptors
    for transition-metal bond chemistry it never saw during training.
  \item \textbf{H3 -- Reactivity (12{,}507 structures).}
    Composition-stratified subsample of three reactivity-flavored
    verticals: \texttt{tm\_react} (5k), \texttt{electrolytes\_reactivity}
    (5k), \texttt{pmechdb} (2.5k). Per-vertical formulas are chosen by
    deterministic BLAKE2b hash so a reader can reproduce membership
    without rerunning the sampler. Tests transfer to non-equilibrium and
    transition-state-adjacent geometries.
  \item \textbf{H6 -- Lanthanide-ligand pairs (2{,}589 structures).}
    Same construction as H1 applied to (Ln, partner) bonds; shape
    $(3,2,2,1,0)$, \texttt{seed=190}; full procedure in
    Appendix~\ref{app:heldout-h6}. The set is smaller than H1 because the
    Ln-bonded subset is itself smaller (66{,}677 structures vs
    651{,}742) and the highest-frequency band is empty.
  \item \textbf{H7 -- Large systems (18{,}200 structures).} Strict
    cutoff \texttt{n\_atoms > 250}. Inherits OMol25's large-system
    stress axis unchanged. Targets size extrapolation in the regime
    where several descriptor methods (notably fuzzy integration and
    grid-based QTAIM) are most expensive, and where the per-atom
    signal-to-noise ratio in DFT outputs is hardest to maintain.
  \item \textbf{H8 -- Large net charges (12{,}393 structures).} Strict
    cutoff \texttt{net\_charge\_abs > 4} (equivalently $|q|\geq 5$).
    Targets the high-charge tail where partial-charge schemes disagree
    the most (cf.\ \S\ref{sec:noise}) and where SCF behavior is most
    sensitive to functional and basis choice.
\end{itemize}

We deliberately use pair-level held-outs (H1, H6) rather than
whole-element held-outs because the goal of this benchmark is to produce
models that are usable across the periodic table, not models that have a
hole where transition metals or lanthanides should be. Dropping every
\texttt{has\_tm} structure would remove 17.24\% of the dataset and evict
whole verticals (\texttt{low\_spin\_23}, \texttt{tm\_react},
\texttt{ml\_mo}) from training entirely; dropping every
\texttt{has\_lanthanide} structure would remove 1.78\% and similarly
empty the Ln-bearing verticals. In both cases the resulting model would
simply have no signal to predict on for those elements at inference
time. Holding out specific (metal, ligand) pairs keeps every element
represented in training while still measuring rare-chemistry transfer,
which is the construction the OMol25 paper uses for the same reason.

\subsection{Holdout-Aware Splitting Pipeline}
\label{sec:eval-pipeline}

\begin{itemize}
  \item \textbf{Step 1 -- Pull holdouts.}
    \texttt{pull\_holdout\_records.py} reads each holdout CSV,
    materializes per-set LMDBs under \texttt{data/holdouts/<set>/}, and
    emits a \texttt{manifest\_holdout.parquet} keyed on the LMDB key
    convention
    \texttt{relpath(folder, root\_dir).replace(os.sep, "\_\_")}.
  \item \textbf{Step 2 -- Exclude.}
    \texttt{multi\_vertical.load\_exclusion\_set} re-derives LMDB keys
    from the manifest \texttt{rel\_path} column and unions
    H1/H3/H6/H7/H8 into one per-vertical exclusion set.
  \item \textbf{Step 3 -- Split.} \texttt{split\_descriptor\_lmdbs.py}
    reads \texttt{structure.lmdb}, drops excluded keys, assigns each
    remaining key to train/val/test by
    \texttt{BLAKE2b(formula\_hill) \% 10}, then partitions each of the
    8 descriptor LMDBs (\texttt{structure}, \texttt{charge},
    \texttt{qtaim}, \texttt{bond}, \texttt{fuzzy}, \texttt{other},
    \texttt{orca}, \texttt{timings}) into the assigned split.
  \item Output layout:
    \texttt{<output\_dir>/<vertical>/<split>/<descriptor>.lmdb}.
  \item HPC parallelism: \texttt{--include\_verticals X} +
    \texttt{--merge\_from DIR}, mirroring the pull pipeline.
    Per-vertical \texttt{split\_assignment.json} is the durable,
    race-free record (no global plan file).
  \item Lock-step audit: across all 8 descriptors, source row count $=$
    train $+$ val $+$ test $+$ excluded for every vertical (verified on
    \texttt{ani2x}: 215{,}642 source $=$ 168{,}783 + 17{,}417 +
    29{,}442, $n_{\mathrm{excluded}}=0$).
\end{itemize}

\subsection{Metrics}
\label{sec:eval-metrics}

\begin{itemize}
  \item Per task: MAE, RMSE, Pearson $r$.
  \item \textbf{T1} stratifications: per-method, per-element,
    per-vertical, per-charge-bucket.
  \item \textbf{T2} stratifications: per-element-pair, per-bond-type
    (covalent / metal-ligand / H-bond), per-vertical.
  \item \textbf{T3} stratifications: macro-F1, per-element-pair
    confusion, per-distance-bucket recall.
  \item Held-out sets: same metrics, plus the gap to the corresponding
    main-test number.
\end{itemize}

\subsection{Methodology and Reporting}
\label{sec:eval-method}

\begin{itemize}
  \item Train on the 8 train buckets. Tune on val (bucket 8). Final
    numbers reported once on test (bucket 9) and on each held-out set,
    with no further tuning.
  \item Held-out sets must not be used for hyperparameter selection or
    early stopping.
  \item Reference \texttt{score-predictions} CLI deferred to v2:
    shipping it for v1 without a calibrated reference model would expose
    the protocol to garbage-in adoption. v2 will release the script
    alongside a reference model whose error sits between the
    cross-method noise floor and naive baselines.
  \item Cross-method noise floors from \S\ref{sec:noise} are a lower
    bound on achievable error: any model below the floor is overfitting
    noise, not learning chemistry. Future work using this benchmark
    should report the floor alongside model error.
  \item Per-vertical post-split distribution: a transparency artifact
    (notebook + appendix table) listing realized train/val/test counts
    per vertical so readers can weight conclusions by vertical-level
    coverage.
\end{itemize}

\subsection{Optional Floor Baseline}
\label{sec:eval-floor}

\begin{itemize}
  \item \textbf{D4} -- per-element-pair lookup table emitting the median
    descriptor value for the matching pair from the train bucket.
  \item Submitted as a floor, never as a baseline. Page-budget
    contingent.
\end{itemize}

